Dish: Distribution-oblivious Shell Scripting

### 2019-11-27 pldi/draft.tex
Polish technical report

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Distributed systems offer notable benefits over centralized ones.
  Reaping these benefits, however, requires programming in a way that makes distribution explicit---\eg via new programming languages, distributed frameworks, or source-level annotations.
  This paper presents \sys, a shell variant that automatically and correctly scales out \unix shell pipelines. 
  Its insight is that pipelines already express stream computations that can be automatically distributed.
  To distribute them, \sys
    decomposes primitives into distributability classes,
    identifies high-distributability stages,
    applies rewriting rules for largest possible subprograms,
    and orchestrates their execution.
Experiments with complex pipelines show substantial speedups and the ability to operate on large input datasets, all without any developer input.
\end{abstract}

@@

Distributed systems offer significant benefits over their centralized counterparts---for example, they can speed up expensive computations or can process data that would not fit into any single machine.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
% Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
% The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.
Despite these benefits, their development remains different from and significantly more difficult than their centralized counterparts.
Whereas anyone can quickly stitch together a Bash script to compute on a single computer, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple ones requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{taurus:14, dios:13, andromeda:15, pywren:17, futuredata:18, nefele:18}.
% 
To understand this sharp contrast, consider a Bash pipeline calculating term frequencies over a set of inputs:

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}
\smallskip
\begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
cat * | tr -cs A-Za-z\n | tr A-Z a-z |    ($$(p_1)$$)
  sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}
\smallskip
% \noindent
Program $p_1$ creates a character stream, breaks it into words, transliterates to lower-case, sorts to group duplicates, reduces duplicates to a counter, sorts counters in descending order, picks the top five results, and writes them to a file \ttt{out}.
Combining several features, \unix makes small tasks easy to express;
  for a user with \emph{one} computer and a small input size, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

Unfortunately, developing $p_1$'s distributed equivalent requires significant effort.
% For a user with many computers and larger inputs, 
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
For simple pipelines, fitting into restricted models of computation, this effort amounts to expressing the computation using the primitives provided by a big-data framework~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} or domain-specific language~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp}---an unjustifiable cost for small, one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, would involve a full-fledged distributed programming language~\cite{erlang:96, lopes1997d, acute:05, mace:07, cloudhaskell:11, ScalaLoci:18}. %  or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In both cases, manual rewriting is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality.
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

To answer this question we develop \sys, a shell variant that transforms traditional \unix pipelines into their distributed equivalents, enabling \emph{distribution-oblivious programming}.
The key insight behind \sys is that a large part of the shell language can be encoded in a distributed dataflow graph model, providing structure to distribute computations.
%% \kk{Replaced that: the language of the Unix shell already encodes stream processing, providing most of the information required to distribute a computation. }
% FIXME \kk{I am not sure whether stream processing provides information for distributing a computation on its own. Maybe we can say that the shell exposes operator level parallelism, and that this is the first step to distributing a computation}.
\sys builds on this insight with a careful study of the distributability properties of shell primitives and commands, a definition of a dataflow graph model that is suitable for the shell, a set of parallelization optimizations that preserve the behaviour of the graph, and a late-bound command-prefixing scheme for deferring scheduling decisions to the distributed planner (\ie when runtime information is available).
A \sys-enabled $p_1$ will run \ttt{tr}s on parallel streams,
  use a mostly-parallel tree of \ttt{sort}s,
  run \ttt{uniq} mostly in parallel,
  execute \ttt{head} in one of the streams,
  and (trivially) merge streams in \ttt{out}.

The combination of automated transformations for program distribution with the ability to maintain correct non-distributed semantics results in several benefits.
% \sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
First, shell users composing pipelines (or simply running legacy pipelines on massive datasets) can see scalability benefits without any manual effort---no need for \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}, or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.
Second, developers of new shell commands can use a lightweight domain-specific language to express their distributability properties rather than ad-hoc, command-specific flags such as {\tt -t},  {\tt NUM\_THREADS}, \ttt{-t}, \ttt{-p} \etc
Most importantly, \sys provides an architectural lesson for system designers---namely, that large-scale efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent~\cite{ousterhout1988sprite, mullender1990amoeba, pike1990plan9, barak1998mosix} would have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.
% Nice critique on parallel: https://bugs.debian.org/cgi-bin/bugreport.cgi?bug=597050#75

@@

The paper is structured as follows.
It starts by introducing the necessary background on shell scripting and overviewing \sys~\sx{bg}.
Sections \ref{distributability}--\ref{impl} highlight key
contributions:
\begin{itemize}

  \item
  \S\ref{distributability} studies several distributability classes and introduces the distributability characterization language.

  \item
  \S\ref{ir} presents a dataflow graph model for encoding shell pipelines, and a set of parallelization transformations that preserve the semantics of the sequential program.

  \item
    \S\ref{impl} describes the \sys implementation and addresses several challenges related to the translation and orchestration of distributed shell programs.
    %% details other concerns, such as distributed state management and extensibility. 
\end{itemize}

\noindent
\sys's evaluation~\sx{eval} shows significant speedups (4--10$\times$) for parallel execution using a combination of real pipelines. %micro-bench\-marks %---popular shell one-liners that highlight certain features---and a multi-line macro-benchmark---a popular weather analysis script.
After a comparison with related prior work~\sx{related}, the paper closes with a discussion of limitations and possible future directions~\sx{discussion}.
% \bigskip
% \begin{quote}
% \footnotesize
Additional material, in anonymized form, is available at
\href{https://git.io/Je6Nk}{https://git.io/Je6Nk}.
% \end{quote}

@@

\heading{Command-line Options}
As described earlier~\sx{bg}, com\-mand-line options and environment variables can affect the operation of a command.
Options are a particularly popular way to control a command, 
  directly affecting a command's distributability classification.
That is, commands are assigned to default classes, and options can change the class in which they belong.

@@

It assumes that the arguments of a command can always be separated to
options and file arguments (checking if they start with a
\ttt{-}~\footnote{Note that the single character \ttt{-} is not an
  option as it is most often used to designate that a command reads
  from its standard input}). A file written in the language contains a
record containing a mapping from predicates to assignments for each
command. Predicates can be simple boolean formulas on the existence
and the value of any option. Each assignment contains the command
category, the sequence of its input sources, and its output
source. Inputs to commands are a sequence of file arguments and its
standard input. Similarly, the output is either its standard output or
a file argument. An example record for comm can be seen below:

@@

\tr{We don't need to designate the stderror, because we assume that it
  is never the main output of a command and it will never be used by
  the input of another command in the pipeline. If someone indeed
  wants to do this, they can just use a redirect I think to get around
  it.}

\tr{If we talk about the language it would be good to give some
  statistics about how many commands (out of the ones that we have
  classified) can be represented with one, two clauses etc. If we have
  a different solution (like python function that given a command and
  its arguments returns the category) then we could talk about how
  many lines of code it took to categorize all the commands that we
  did.}

\tr{Note: Since a categorization language will probably not be
  complete (especially the one showing the input arguments) there
  should be a backup mechanism, when the language is not expressive
  enough, like a python function that categorizes the command, and
  returns its input argument if it is stateless.}

\tr{(Maybe) We should have a crisp point about why we have this categorization
  language, and why don't we just allow someone to write a function in
  python for each command, that given the command and its arguments,
  returns whether the command category. Possible arguments include,
  the fact that this language is simpler to use, especially by non
  experts that just run some script but have installed some commands
  that are not ``supported''. Another possible argument is that we
  could be able to reason about the constructs of the categorization
  lagnauge, and that they will be easier to read and
  understand. Another argument is that since these categorizations
  should be shareable among users, it would be bad to execute
  arbitrary python code, so using this language, expressivity is
  limited. All of these arguments are a little bit weak though. Niko,
  what do you think?}

%% \heading{Pure Commands}
%
\noindent
Together with the commands in \sta, that are always easy to
parallelize, there also exists a subclass of commands in \pur that can
be parallelized by separating them into a map and a reduce
operation. In order to allow command developers to extend \sys with
information about how to parallelize new pure commands, \sys offers a
programmatic interface to specify how to distribute a pure command.

\tr{I am not sure a general interface is so easy to design. It needs
  more though. It might be beneficial to just talk about sort and wc
  here and how we implemented them and nothing more. Or maybe this
  could then go to the implementation? Or maybe say that one can write
  a python function that given a node of the graph, transforms it into
  many. I am not sure what is best...}

@@

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable
representation. As mentioned in \Cref{bg:pipelines}, the two main
abstractions of the shell, are files that contain data, and commands
that communicate through these files. We propose a slight variation of
the standard dataflow graph model, that is commonly found in stream
processing systems, as the representation of a distributed shell
computation. The main benefit of the dataflow graph model is that it
clearly exposes task parallelism, as different nodes can be
independent processing workers. In our case each node represents a
command, and each edge represents a file.

\subsection{Graph Components}

@@

As mentioned above, the dataflow graph exposes task
parallelism. Different nodes in the graph can be executed
concurrently, communicating through the edges.
%% Task level parallelism is already exposed by the shell, by using
%% pipes \ttt{|} and the background operator \ttt{\&}.
Unfortunately, simple task parallelism leaves a lot to be desired,
since scaling a computation requires manually introducing more
operators.

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. Exploiting data parallelism is
one of the major reasons that popular distributed stream processing
systems (such as MapReduce \cite{mapreduce:08} and Apache Spark
\cite{spark:12}) achieve significant performance. Unfortunately, the
optimizations in many of these systems do not preserve the dataflow
model semantics, blurring the lines between specification,
optimization, and implementation.

In contrast, our dataflow model inherently supports the data
parallelism that is found in shell programs. More precisely, data
parallelism can be easily exposed as task parallelism in our model. A
very important consequence of this is that we can iteratively apply
optimizations that expose data parallelism, or even revert them when
the resulting dataflow graph is too wide. This flexibility unifies the
unoptimized and optimized versions of a computation, enabling thinking
about them with the same conceptual tools.

@@

\tr{This is a nice characterization, which I think is true in most
  settings. There is a caveat though: You can have stateless programs
  that emit output when they start (before they consume the stream)
  and/or when they consume the EOF symbol. In this case I am not sure
  the equation above holds.}

Similarly, it was mentioned that some pure commands (such as
\ttt{sort} and \ttt{wc}) can be parallelized using divide-and-conquer
techniques. More formally, there pure commands $f$ can be implemented
as a map $m$ and a reduce $r$ function, satisfying the following:

@@

@@

As mentioned in \Cref{ir}, in order to generate the distributed
implementation of a shell script, \ttt{Dish} first translates it to
a dataflow graph. However, there are several components in shell
programs that cannot be arbitrarily parallelized without affecting the
program behaviour, e.g. commands producing an irreversible effect, or
commands connected with a \ttt{\&\&} operator. In addition, the
fact that the shell is highly dynamic, further increases the
challenge of correctly identifying what parts can be distributed.

@@

\subsection{Distributable Regions}

%% \kk{The goal of this subsection is to explain that Dish does a safe
%%   but effective analysis to identify which parts of the scripts to
%%   parallelize.}

Consider the following example:

@@

Concretely, a distributable region cannot permeate through barrier
constructs (e.g. \ttt{;}, \ttt{\&\&}, and function definitions). On
the other hand, the pipe operator \ttt{|} and the parallel composition
operator \ttt{\&} enable parallelization, and extend distributable
regions. Intuitively, distributed regions correspond to subsections of
the program that would be allowed to execute concurrently by different
processes in the POSIX standard~\cite{posix}.

@@

\subsection{Translation Pass}

Dish performs a depth first search pass on the AST of the given shell
program, extending the distributable regions from the bottom up,
translating their independent components to dataflow graph
nodes, until a barrier construct is reached. Note that barrier
constructs can exist in a dataflow graph, as long as they don't
extend beyong a single node. An illustrative example follows:

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
  (sort f1 > g1 && sort f2 > g2) &
  sort -m g1 g2
\end{lstlisting}

In the above example, while \ttt{sort f1 > g1}, \ttt{sort f2 > g2},
and \ttt{sort -m g1 g2} are all in different dataflow graphs, the
dataflow graph that corresponds to the whole program, has two nodes,
\ttt{sort f1 > g1 \&\& sort f2 > g2} and \ttt{sort -m g1 g2}.

In order to identify possible opportunities for parallelism, the
translation pass also identifies the category of each node, together
with its input and output, using records written by developers using
the language that was described in \Cref{ext}. Some interesting cases
of the analysis algorithm are shown in pseudocode below:

@@

\subsection{Dish Optimizer}

@@

% \subsection{Planning Activators and Just-in-Time Planning}
% 
% \nv{Nikos working on this and next}
% 
% % \kk{This whole section for just in itme planning, is also relevant for
% %   the front end. If the front end is executed just in time, it would
% %   have more information, and maybe it could make a sound analysis.}
% 
% Important point: Shell is extremely dynamic. Because of this, a
% distributor cannot decide how to distribute a subexpression
% statically, by just reading the script, (as in many other systems like
% MapReduce, Spark, etc), since most of the information is not there
% before the script starts executing. Environment variables,
% unexpanded(unevaluated) strings \kk{Make sure that terminology is
%   consistent with Greenberg}, could all contain information that is
% valuable to make the distribution plan \kk{give an example}.
% 
% 
% Future work on this: calling the shell shtepper to only partially
% evaluate strings, just expand arguments, and not do any significant
% computation. Even better, we could be calling the shtepper just after
% having parsed the ast, and decide on maximal distributable subtrees as
% late as possible in the process. This would be awesome.
% 
\subsection{Merge and Split Implementation}

It is common in the dataflow graph for a node to have many incoming
edges, or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might only read their input from one
source (e.g. stdin). In this case, the distributed implementation
needs to concatenate input files (or conversely split output
files).
%
%% \tr{Show an example graph! of cat | tr | sort. The command in the
%%   middle must only take input from one file (maybe just stdin).}
%
We address this issue using primitive shell constructs for file
manipulation. More concretely a merger can be implemented by a simple
\verb|cat|. Given a set of input files \verb|in1, in2, ...| and an
output file \verb|out|, a merger can be implemented as:

@@

% The first interesting result 
The effort required to write this pipeline is astonishingly low:\footnote{
  Some effort is required to understand NCDC's weather format, but this is true for any program processing this dataset.
}
  it amounts to 14 stages and, when expressed as a single line, only 240 characters.
The Hadoop program, expressed in Java and only capturing the 4 last stages, is about 137 lines of Java code---not accounting for code moving input and output in and out of HDFS.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Distributed systems offer notable benefits over centralized ones.
  Reaping these benefits, however, requires programming in a way that makes distribution explicit---\eg via new programming languages, distributed frameworks, or source-level annotations.
  This paper presents \sys, a shell variant that automatically and correctly scales out distribution-oblivious shell pipelines. 
  Its insight is that pipelines already express stream computations that can be automatically distributed.
  To distribute them, \sys
    decomposes primitives into distributability classes,
    identifies high-distributability stages,
    applies rewriting rules for largest possible subprograms,
    and orchestrates the execution of the distributed program.
Experiments for complex pipelines show substantial speedups and the ability to operate on large input datasets, all without any developer input.
\end{abstract}

@@

Distributed systems offer significant benefits over their centralized counterparts---for example, they can speed up expensive computations or can process data that would not fit into any single machine.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
% Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
% The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.
Despite these benefits, their development remains different from and significantly more difficult than their centralized counterparts.
Whereas anyone can quickly stitch together a Bash script to compute on a single computer, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple ones requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{taurus:14, dios:13, andromeda:15, pywren:17, futuredata:18, nefele:18}.

To understand this sharp contrast, consider a Bash pipeline calculating term frequencies over a set of inputs:

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}
\smallskip
\begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
cat * | tr -cs A-Za-z\n | tr A-Z a-z |    ($$[p_1]$$)
  sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}
\smallskip
% \noindent
Program $p_1$ creates a character stream, breaks it into words, transliterates to lower-case, sorts to group duplicates, reduces duplicates to a counter, sorts counters in descending order, picks the top five results, and writes them to a file \ttt{out}.
Combining several features, \unix makes small tasks easy to express;
  for a user with \emph{one} computer and small input size, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

Unfortunately, developing $p_1$'s distributed equivalent requires a significant effort.
% For a user with many computers and larger inputs, 
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
For simple pipelines, fitting into restricted models of computation, this effort amounts to expressing the computation using the primitives provided by a big-data framework~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} or domain-specific language~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp}---an unjustifiable cost for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, would involve a full-fledged distributed programming language~\cite{erlang:96, lopes1997d, acute:05, mace:07, cloudhaskell:11, ScalaLoci:18}. %  or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In both cases, manual rewriting is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality.
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

To answer this question, this paper presents \sys, a shell variant that transforms traditional shell pipelines into their distributed equivalents, enabling \emph{distribution-oblivious programming}.
The key insight behind \sys is that a large part of the shell language can be encoded using a dataflow graph model, providing structure to distribute computations.
%% \kk{Replaced that: the language of the Unix shell already encodes stream processing, providing most of the information required to distribute a computation. }
% FIXME \kk{I am not sure whether stream processing provides information for distributing a computation on its own. Maybe we can say that the shell exposes operator level parallelism, and that this is the first step to distributing a computation}.
\sys builds on this insight with a careful study of the distributability properties of shell primitives and commands, a definition of a dataflow graph model that is suitable for the shell, a set of parallelization optimizations that preserve the behaviour of the graph, and a late-bound command-prefixing scheme for deferring scheduling decisions to the distributed planner (\ie when runtime information is available).
A \sys-enabled $p_1$ will run \ttt{tr}s in parallel streams,
  use a mostly-parallel tree of \ttt{sort}s,
  run \ttt{uniq} mostly in parallel,
  \ttt{head} in one of the streams,
  and (trivially) merge streams in \ttt{out}.

The combination of automated transformations for program distribution with the ability to maintain correct non-distributed semantics results in several benefits.
% \sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
First, shell users composing pipelines (or simply running legacy pipelines on massive datasets) can see scalability benefits without any manual effort---no need for \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}, or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.
Second, developers of new shell commands can use a concise a domain-specific language to express their distributability properties rather than using ad-hoc, command-specific flags such as {\tt -t},  {\tt NUM\_THREADS}, \ttt{-t}, \ttt{-p} \etc
Most importantly, \sys provides an architectural lesson for system designers---namely, that large-scale research efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent~\cite{ousterhout1988sprite, mullender1990amoeba, pike1990plan9, barak1998mosix} would have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.
% Nice critique on parallel: https://bugs.debian.org/cgi-bin/bugreport.cgi?bug=597050#75

@@

The paper is structured as follows.  It starts with a background
section outlining pipeline concepts and overviewing \sys~\sx{bg}.
Sections \ref{distributability}--\ref{impl} highlight key
contributions:
\begin{itemize}

  \item
  \S\ref{distributability} describes several scalability and distributability classes and introduces the distributability characterization language.

  \item
  \S\ref{ir} presents a dataflow graph model that can encode shell programs, and a set of parallelization transformations that preserve its semantics.

  \item
    \S\ref{impl} describes the \sys implementation and addresses several challenges related to the translation and orchestration of shell programs.
    %% details other concerns, such as distributed state management and extensibility. 
\end{itemize}

\noindent
\sys's evaluation~\sx{eval} shows significant speedups on both parallel and distributed execution, using a combination of micro-bench\-marks---popular shell one-liners that highlight certain features---and a multi-line macro-benchmark---a popular weather analysis script.
After a comparison with related prior work~\sx{related}, the paper closes with a discussion of limitations and possible future directions~\sx{discussion}.
% \bigskip
% \begin{quote}
% \footnotesize
Additional material, in anonymized form, is available at
\href{https://git.io/Je6Nk}{https://git.io/Je6Nk}.
% \end{quote}

@@

\heading{Command-line Options}
As described earlier~\sx{bg}, command-line options and environment variables can affect the operation of a command.
Options are a particularly popular way to control a command, 
  directly affecting a command's distributability classification.
That is, commands are assigned to default classes, and options can change the class in which they belong.

@@

It assumes that the arguments of a command can always be separated to
options and file arguments (checking if they start with a
\ttt{-}~\footnote{Note that the single character \ttt{-} is not an
  option as it is most often used to designate that a command reads
  from its standard input}). A file written in the language contains a
record containing a mapping from predicates to assignments for each
command. Predicates can be simple boolean formulas on the existence
and the value of any option. Each assignment contains the command
category, the sequence of its input sources, and its output
source. Inputs to commands are a sequence of file arguments and its
standard input. Similarly, the output is either its standard output or
a file argument. An example record for \ttt{comm} can be seen below:

@@

\tr{\kk{We don't need to designate the stderror, because we assume that it
  is never the main output of a command and it will never be used by
  the input of another command in the pipeline. If someone indeed
  wants to do this, they can just use a redirect I think to get around
  it.}}

\tr{\kk{If we talk about the language it would be good to give some
  statistics about how many commands (out of the ones that we have
  classified) can be represented with one, two clauses etc. If we have
  a different solution (like python function that given a command and
  its arguments returns the category) then we could talk about how
  many lines of code it took to categorize all the commands that we
  did.}}

\tr{\kk{Note: Since a categorization language will probably not be
  complete (especially the one showing the input arguments) there
  should be a backup mechanism, when the language is not expressive
  enough, like a python function that categorizes the command, and
  returns its input argument if it is stateless.}}

\tr{\kk{(Maybe) We should have a crisp point about why we have this categorization
  language, and why don't we just allow someone to write a function in
  python for each command, that given the command and its arguments,
  returns whether the command category. Possible arguments include,
  the fact that this language is simpler to use, especially by non
  experts that just run some script but have installed some commands
  that are not ``supported''. Another possible argument is that we
  could be able to reason about the constructs of the categorization
  lagnauge, and that they will be easier to read and
  understand. Another argument is that since these categorizations
  should be shareable among users, it would be bad to execute
  arbitrary python code, so using this language, expressivity is
  limited. All of these arguments are a little bit weak though. Niko,
  what do you think?}}

%% \heading{Pure Commands}
%
\noindent
In addition to the commands in \sta that can be parallelized in a
straightforward way, there also exists a subclass of commands in \pur
that can be parallelized using divide-and-conquer techniques (as
mentioned in \cref{cmd}). In order to support the parallelization of
arbitrary commands in this category, \sys allows command developers to
provide a custom map and reduce step for these commands, where the map
is in \sta and the reduce in \pur, and their composition produces the
same output with the initial command. \sys can then replace occurences
of these commands with their map and reduce parts, enabling their
parallelization.

\tr{\kk{I am not sure a general interface is so easy to design. It needs
  more though. It might be beneficial to just talk about sort and wc
  here and how we implemented them and nothing more. Or maybe this
  could then go to the implementation? Or maybe say that one can write
  a python function that given a node of the graph, transforms it into
  many. I am not sure what is best...}}

@@

\kk{Move some variation of that to intro/overview:

In the core of \sys there is a dataflow graph model that is inspired
by the ones that are used in popular distributed stream processing
systems (such as MapReduce \cite{mapreduce:08} and Apache Spark
\cite{spark:12}). These systems exploit data parallelism by
parallelizing nodes of the dataflow graph. Unfortunately, the
optimizations in many of these systems do not preserve the dataflow
model semantics, blurring the lines between specification,
optimization, and implementation. In contrast, our dataflow model
inherently supports the data parallelism that is found in shell
programs. More precisely, data parallelism can be easily exposed as
task parallelism in our model. We exploit that by developing a set of
parallelization-exposing optimizations that preserve the dataflow
model semantics, and can be represented as graph transformations. The
\sys optimizer then applies these transformations in an iterative way,
optimizing the program to exploit the available computational
resources.

%% The standard dataflow graph exposes task parallelism. Different nodes
%% in the graph can be executed concurrently, communicating through the
%% edges.
%% Task level parallelism is already exposed by the shell, by using
%% pipes \ttt{|} and the background operator \ttt{\&}.
%% Unfortunately, simple task parallelism leaves a lot to be desired,
%% since scaling a computation requires manually introducing more
%% operators.

%% In order to adequately utilize the available computational resources,
%% a distributed implementation should exploit data parallelism. Data
%% parallelism is based on the observation that some operations produce
%% the same result when performed independently on subsets of the input
%% data, merging the results afterwards. Exploiting data parallelism is
%% one of the major reasons that popular distributed stream processing
%% systems (such as MapReduce \cite{mapreduce:08} and Apache Spark
%% \cite{spark:12}) achieve significant performance.

%% A very important consequence of this is that we can iteratively apply
%% optimizations that expose data parallelism, or even revert them when
%% the resulting dataflow graph is too wide. This flexibility unifies the
%% unoptimized and optimized versions of a computation, enabling thinking
%% about them with the same conceptual tools.
}

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable representation.
In \Cref{graph-components} we propose a slight variation of the
standard dataflow graph model as a candidate intermediate
representation. The main benefit of our model is that it exposes both
task and data parallelism. In \Cref{ir:transformations} we define a
set of optimizations that expose and exploit data parallelism of
commands in \sta and \pur as semantics-preserving graph
transformations.

\subsection{Graph Components}
\label{graph-components}

As mentioned in \Cref{bg:pipelines}, the two main abstractions of the
shell are files that contain data, and commands that communicate
through these files. Based on that, we define nodes in our graph model
to represent commands, and edges to represent files.

@@

@@

%% This is a nice characterization, which I think is true in most
%% settings. There is a caveat though: You can have stateless programs
%% that emit output when they start (before they consume the stream)
%% and/or when they consume the EOF symbol. In this case I am not sure
%% the equation above holds.

\noindent
This means that we get the same output by applying a command $f$ to a
concatenation of two inputs $x, x'$, or by concatenating the result
produced by the application of $f$ to these inputs
separately. Similarly, it was mentioned that some pure commands (such
as \ttt{sort} and \ttt{wc}) can be parallelized using
divide-and-conquer techniques. More formally, there pure commands $f$
can be implemented as a map $m$ and a reduce $r$ function, satisfying
the following:

@@

\noindent
Similarly to the above, this means that we get the same output by
applying $f$ to a concatenation of two inputs $x, x'$, or by applying
the reduce function $r$ to the concatenation of the outputs produced
by applying map $m$ to each of these inputs.

@@

In this section, we describe challenges related to the implementation
of \sys. More concretely, in \Cref{front-end} we describe how the \sys
front-end translates a shell script to the dataflow graph that was
defined in \Cref{ir}, and in \Cref{optimizer}, we describe how the
\sys optimizer distributes and implements a given dataflow graph.

\subsection{Front End}
\label{front-end}

As mentioned in \Cref{ir}, in order to generate the distributed
implementation of a shell script, \ttt{Dish} first translates it to a
dataflow graph. However, there are several components in shell
programs that cannot be arbitrarily parallelized without affecting the
program behaviour, e.g. commands connected with a \ttt{\&\&}
operator. In addition, the fact that the shell is highly dynamic,
further increases the challenge of correctly identifying what parts
can be distributed, e.g. by not being able to know what file does an
unexpanded command argument refer to.

@@

\heading{Distributable Regions}
%
%% \kk{The goal of this subsection is to explain that Dish does a safe
%%   but effective analysis to identify which parts of the scripts to
%%   parallelize.}
%
Consider the following example:

@@

Concretely, a distributable region cannot permeate through barrier
constructs (e.g. \ttt{;}, \ttt{\&\&}, and function definitions). On
the other hand, the pipe operator \ttt{|} and the parallel composition
operator \ttt{\&} enable parallelization, and extend distributable
regions. Intuitively, distributed regions correspond to subsections of
the program that would be allowed to execute concurrently by different
processes in the POSIX standard~\cite{posix}. Consider the following
illustrative example:

@@

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
  (tr ... f1 > g1 && tr ... f2 >> g1) &
  grep ``foo'' g1
\end{lstlisting}

In the above example, while \ttt{tr ... f1 > g1}, \ttt{tr ... f2 >>
  g1}, and \ttt{grep ``foo'' g1} are all in different dataflow graphs,
the dataflow graph that corresponds to the whole program, has two
nodes, \ttt{tr ... f1 > g1 \&\& tr ... f2 >> g1} and \ttt{grep ``foo''
  g1}.

\heading{Translation Pass}
%
Dish performs a depth first search pass on the AST of the given shell
program, extending the distributable regions from the bottom up,
translating their independent components to dataflow graph nodes,
until a barrier construct is reached. In order to identify possible
opportunities for parallelism, the translation pass also identifies
the category of each node, together with its input and output, using
records written by developers using the language that was described in
\Cref{ext}. Some interesting cases of the analysis algorithm are shown
in pseudocode below:

@@

\subsection{Optimizer}
\label{optimizer}

@@

% \subsection{Planning Activators and Just-in-Time Planning}
% 
% \nv{Nikos working on this and next}
% 
% % \kk{This whole section for just in itme planning, is also relevant for
% %   the front end. If the front end is executed just in time, it would
% %   have more information, and maybe it could make a sound analysis.}
% 
% Important point: Shell is extremely dynamic. Because of this, a
% distributor cannot decide how to distribute a subexpression
% statically, by just reading the script, (as in many other systems like
% MapReduce, Spark, etc), since most of the information is not there
% before the script starts executing. Environment variables,
% unexpanded(unevaluated) strings \kk{Make sure that terminology is
%   consistent with Greenberg}, could all contain information that is
% valuable to make the distribution plan \kk{give an example}.
% 
% 
% Future work on this: calling the shell shtepper to only partially
% evaluate strings, just expand arguments, and not do any significant
% computation. Even better, we could be calling the shtepper just after
% having parsed the ast, and decide on maximal distributable subtrees as
% late as possible in the process. This would be awesome.
%

\heading{Merge and Split Implementation}
%
It is common in the dataflow graph for a node to have many incoming
edges, or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might only read their input from one
source (e.g. stdin). In this case, the distributed implementation
needs to concatenate input files (or conversely split output
files).
%
%% \tr{Show an example graph! of cat | tr | sort. The command in the
%%   middle must only take input from one file (maybe just stdin).}
%
We address this issue using primitive shell constructs for file
manipulation. More concretely a merger can be implemented by a simple
\verb|cat|. Given a set of input files \verb|in1, in2, ...| and an
output file \verb|out|, a merger can be implemented as:

@@

% The first interesting result 
The effort required to write this pipeline is astonishingly low:\footnote{
  Some effort is required to understand NCDC's weather format, but this is true for any program processing this dataset.
}
  it amounts to 14 stages and, when expressed as a single line, only 240 characters.
The Hadoop program, expressed in Java and only capturing the 4 last stages, is about 117 lines of Java code---not accounting for code moving input and output in and out of HDFS.

In terms of execution time, with two Hadoop datanodes running on Hadoop 3.2.1 and Openjdk 11.0.5 without any replication (\ie fault-tolerance), the Hadoop program takes 6m22s for data processing and 1:09 for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---but this includes downloading and uncompressing the data files.
